---
title: "Adaptive Method"
author: "Carlos V. Paradis"
date: "April 19, 2018"
output: 
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---


```{r messages = FALSE, warning = FALSE}
s <- suppressPackageStartupMessages
s(library(zoo))
s(library(data.table))
s(library(lubridate))
s(library(knitr))
s(library(ggplot2))
```

# Introduction

This notebook uses the outdoor temperature, indoor temperature and indoor airspeed to calculate:

 * Outdoor Rolling Average over last 14 days
 * Upper and Lower Bound of Comfort Level 
 * Number of Indoor Temperature Degrees Off the Upper or Lower Comfort Bound 

```{r}
# Parameters
n_hours_in_every_step <- 366 #Rolling mean will be based of the previous 14 days measurements
```

# Method

## Perform Average by Hour

The first step to the code is to construct the outdoor rolling mean. In order to do that, first the outdoor hourly measurement is averaged in hours.

```{r}
# Load and rename columns for consistency -- no space and no upper case column names
outdoor <- fread("~/Github/erdl/thermal_comfort/adaptive_method/data/houses_outdoor_temperature.csv")
colnames(outdoor) <- c("timestamp","outdoor_temperature_f")

# Identify the right types for each column
outdoor$timestamp <- ymd_hms(outdoor$timestamp,tz="HST")
outdoor$outdoor_temperature_f <- as.numeric(outdoor$outdoor_temperature_f)
outdoor$ymd_h <- outdoor$timestamp

#Ensure timestamps are ordered. Database does not guarantee that.
outdoor <- outdoor[order(timestamp)]

# Reset the minute and second to 0, so we can group by ymd_h. 
minute(outdoor$ymd_h) <- 0
second(outdoor$ymd_h) <- 0
outdoor <- outdoor[,.(outdoor_temperature=mean(outdoor_temperature_f)),by="ymd_h"]
kable(head(outdoor))
```


## Perform 14 Days Hourly Rolling Average

Next, the hourly temperature is used to calculate the 14 days temperature backwards. 

```{r}
# Invert table so slide window goes backwards
inverted_outdoor <- outdoor[order(-ymd_h)]

# Calculate rolling mean and reverts back the output from the rolling mean. 
rolling_mean_outdoor_temperature <- rev(rollapply(inverted_outdoor$outdoor_temperature,
                                            width=n_hours_in_every_step,
                                            FUN=mean))

#
outdoor <- outdoor[n_hours_in_every_step:.N]  
outdoor$rolling_mean_outdoor_temperature <- rolling_mean_outdoor_temperature
colnames(outdoor) <- c("timestamp","outdoor_temperature","rolling_mean_outdoor_temperature")

```

## Simulation Data 

To calculate the boundaries and acceptance, besides the outdoor rolling average, we also need the indoor temperature, and air speed. Because the tables are different, i.e. indoor's temperature and airspeed are hourly, wherewas the outdoor was originally at a minute sampling rate, there is a chance after the rolling average is calculated the number of rows will not match between both tables. This script matches the minimum number of rows between them both.

```{r}
# load indoor data
indoor <- fread("~/Github/erdl/thermal_comfort/adaptive_method/data/house2_livingroom_temperature.csv")[,.(reading_timestamp,value)]
colnames(indoor) <- c("timestamp","indoor_temperature")
indoor$timestamp <- ymd_hms(indoor$timestamp,tz="HST")

# Identify the right types for each column
indoor$timestamp <- ymd_hms(indoor$timestamp,tz="HST")
indoor$outdoor_temperature_f <- as.numeric(outdoor$indoor_temperature)
indoor$ymd_h <- indoor$timestamp

#Ensure timestamps are ordered. Database does not guarantee that.
indoor <- indoor[order(timestamp)]

# Reset the minute and second to 0, so we can group by ymd_h. 
minute(indoor$ymd_h) <- 0
second(indoor$ymd_h) <- 0
indoor <- indoor[,.(indoor_temperature=mean(indoor_temperature)),by="ymd_h"]
colnames(indoor) <- c("timestamp","indoor_temperature")
```


```{r}
dt <- merge(outdoor,indoor,by="timestamp")

# Add Rolling Mean Vector to Simulation Data Column-Wise. Notice this is NOT an inner join.
#min_rows <- min(length(rolling_temperature_f_mean),nrow(simulation_data))

#dt <- simulation_data[1:min_rows]
#dt$rolling_temperature_f_mean <- rolling_temperature_f_mean[1:min_rows]
```

## Calculate boundaries

The comfort level boundaries are calculated using the outdoor rolling average adjusted by the indoor airspeed. 

```{r}
calculate_air_speed_adjustment <- function(in_air_speed){
  return(0.0153*in_air_speed+0.4333)
}

dt$upper_bound <- 0.31*dt$rolling_mean_outdoor_temperature + 60.5
dt$upper_bound_0_fpm <- dt$upper_bound + calculate_air_speed_adjustment(0)
dt$upper_bound_120_fpm <- dt$upper_bound + calculate_air_speed_adjustment(120)
dt$upper_bound_200_fpm <- dt$upper_bound + calculate_air_speed_adjustment(200)


dt$lower_bound <- 0.31*dt$rolling_mean_outdoor_temperature + 47.9 
```

## Calculate Acceptance

With the comfort upper and lower bound, we then compare the indoor temperature against them.

```{r}
dt$is_acceptable_80 <- NA_character_
dt[indoor_temperature > upper_bound]$is_acceptable_80 <- "Unacceptable Hot"
dt[indoor_temperature < lower_bound]$is_acceptable_80 <- "Unacceptable Cold"
dt[indoor_temperature > lower_bound & indoor_temperature < upper_bound]$is_acceptable_80 <- "Acceptable"
```

## Calculate Degrees Off

In addition, how far higher or lower than the upper and lower bound respectively can also be calculated.

```{r}
dt$degrees_off <- NA_real_

dt[indoor_temperature > upper_bound]$degrees_off <- dt[indoor_temperature > upper_bound]$indoor_temperature - dt[indoor_temperature > upper_bound]$upper_bound

dt[indoor_temperature < lower_bound]$degrees_off <- dt[indoor_temperature < lower_bound]$indoor_temperature - dt[indoor_temperature < lower_bound]$lower_bound

dt[indoor_temperature > lower_bound & indoor_temperature < upper_bound]$degrees_off <- 0
```

# Output Table

Finally, we the output table containing all the calculated information.

```{r}
kable(head(dt))
```

# Plot

```{r fig.width= 9, fig.height=5}
n_acceptable <- nrow(dt[is_acceptable_80 == "Acceptable"])
n_unacceptable <- nrow(dt[is_acceptable_80 != "Acceptable"])

st <- min(dt$timestamp)
et <- max(dt$timestamp)
p_start_time <- paste0(year(st),"-",month(st),"-",day(st))
p_end_time <- paste0(year(et),"-",month(et),"-",day(et))



p <- ggplot(data = dt) + 
  geom_point(aes(rolling_mean_outdoor_temperature,indoor_temperature,color=is_acceptable_80)) + 
  geom_line(aes(rolling_mean_outdoor_temperature,upper_bound_200_fpm),color="black") + 
  geom_line(aes(rolling_mean_outdoor_temperature,upper_bound_120_fpm),color="black") + 
  geom_line(aes(rolling_mean_outdoor_temperature,upper_bound),color="black") + 
  geom_line(aes(rolling_mean_outdoor_temperature,lower_bound),color="black") + 
  
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$upper_bound_200_fpm), label = "fpm 200", vjust = -1), size = 3) + 
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$upper_bound_120_fpm), label = "fpm 120", vjust = -1), size = 3) + 
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$upper_bound)-0.6, label = "no adj", vjust = -1), size = 3) + 
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$lower_bound), label = "no adj", vjust = -1), size = 3) + 
  
  
  theme_minimal() + 
  scale_colour_manual(values=c("Unacceptable Cold"="#0072B2","Acceptable"="#009E73","Unacceptable Hot"="#D55E00")) +
  ggtitle("Adaptive Method - House 2 Living Room", 
          subtitle = paste0(p_start_time," to ",p_end_time," | Acceptable: ",n_acceptable," ; Unacceptable: ",n_unacceptable)) + 
  ylab("Indoor Temperature (F)") + 
  xlab("Rolling Mean Outdoor Temperature (F)") 

p
```

