---
title: "Adaptive Method"
author: "Carlos V. Paradis"
date: "April 19, 2018"
output: 
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---


```{r messages = FALSE, warning = FALSE}
s <- suppressPackageStartupMessages
s(library(zoo))
s(library(data.table))
s(library(lubridate))
s(library(knitr))
s(library(ggplot2))
s(library(stringr))
s(library(formattable))
```

# Introduction

This notebook uses the outdoor temperature, indoor temperature and indoor airspeed to calculate:

 * Outdoor Rolling Average over last 14 days
 * Upper and Lower Bound of Comfort Level 
 * Number of Indoor Temperature Degrees Off the Upper or Lower Comfort Bound 

```{r}
# Parameters
n_hours_in_every_step <- 366 #Rolling mean will be based of the previous 14 days measurements
```

# How to use

   * Use the code block "Monitored" or "Simulation" to load the associated data and put it in a common format for the rest of the notebook. (Remember to set the unused code block as eval = FALSE and the other with eval = TRUE).
   * Change the title of the plot on the **data_type** variable
   * Change the location of the file from c_view_house_<number> provided by Eileen on the **indoor** variable
   * You do not need to change fpm here! They are shown as lines in this plot for 0, 120, and 200 fpm.
   * Change the **save_file_path_name** variable with the location and file name.csv you want to save the generated plot. 
   * For filtering some data, See the code block for `Time Window Filter` and decide if you want to apply or not for the current run.
   
If you need the table so it can be plotted in Tableau, you can use **dt** data frame variable and save it to a .csv. 
   
That's all! Once the pipeline and plots get more stable, I will code a script out of this.

# Method

## Monitored Data

Use the following code block if the Adaptive Method will be used in **monitored** data.

```{r Monitored, eval = FALSE}
data_type <- "Monitored House 2 Living Room" #used in the plot title
save_path_and_name <- "~/Desktop/c-view-house-2-adaptive-living-room.svg"

# Outdoor

## Perform Hourly Average

# Load and rename columns for consistency -- no space and no upper case column names
outdoor <- fread("~/Github/erdl/thermal_comfort/adaptive_method/data/houses_outdoor_temperature_new_time_window.csv")
colnames(outdoor) <- c("timestamp","outdoor_temperature_f")

# Identify the right types for each column
outdoor$timestamp <- mdy_hms(outdoor$timestamp,tz="HST") #if exported via navicat the date format will be different than database
#outdoor$timestamp <- ymd_hms(outdoor$timestamp,tz="HST")
outdoor$outdoor_temperature_f <- as.numeric(outdoor$outdoor_temperature_f)
outdoor$ymd_h <- outdoor$timestamp

#Ensure timestamps are ordered. Database does not guarantee that.
outdoor <- outdoor[order(timestamp)]

# Reset the minute and second to 0, so we can group by ymd_h. 
minute(outdoor$ymd_h) <- 0
second(outdoor$ymd_h) <- 0
outdoor <- outdoor[,.(outdoor_temperature=mean(outdoor_temperature_f)),by="ymd_h"]


# Indoor Temperature

indoor <- fread("~/Github/erdl/thermal_comfort/adaptive_method/data/monitored/house2/c_view_house_2_adaptive_living_room_new_time_window.csv")

# Correct types, filter NAs
indoor <- indoor[,.(timestamp=mdy_hms(datetime_hst_hrly),indoor_temperature=as.numeric(indoor_temperature))] #if exported via navicat the date format will be different than database
#indoor <- indoor[,.(timestamp=ymd_hms(datetime_hst_hrly),indoor_temperature=as.numeric(indoor_temperature))]
indoor <- indoor[complete.cases(indoor)]
indoor$ymd_h <- indoor$timestamp

#Ensure timestamps are ordered. Database does not guarantee that.
indoor <- indoor[order(timestamp)]

# Reset the minute and second to 0, so we can group by ymd_h. 
minute(indoor$ymd_h) <- 0
second(indoor$ymd_h) <- 0
indoor <- indoor[,.(indoor_temperature=mean(indoor_temperature)),by="ymd_h"]
colnames(indoor) <- c("timestamp","indoor_temperature")
```

## Simulated Data

Use the following code block if the Adaptive Method will be used in **simulated** data.

```{r Simulated, warning = FALSE, eval = TRUE}
data_type <- "Simulated House 2 Run 3 Time Window" #used in the plot title
save_path_and_name <- "~/Desktop/house2-run-3-time-window.png"
# For simulated data, only 1 table, generated by BEopt, is necessary, as it contains both outdoor and indoor simulated data.

dt <- fread("~/Github/erdl/thermal_comfort/adaptive_method/data/simulated/house-2-run-3.csv")
colnames(dt) <- sapply(str_split(colnames(dt)," - "),"[[",2)
dt <- dt[2:.N,.(outdoor_temperature=as.numeric(`Weather|Outdoor Drybulb`), #First row is units
            indoor_temperature=as.numeric(`Living Space|Indoor Temperature`))]

outdoor <- dt$outdoor_temperature

# Add 366 data points to the start of outdoor vector so indoor first 366 points are not lost when calculating rolling mean.
outdoor <- c(tail(outdoor,n=365),outdoor)

# Create fake timestamp id

timestamp <- ymd_hms("1800-01-01 00:00:00") + hours(1:length(outdoor))

outdoor <- data.table(ymd_h=timestamp,outdoor_temperature=outdoor)

# Do the same for temperature:

dt$timestamp <- ymd_hms("1800-01-01 00:00:00") + hours(366:(nrow(dt)+366))

indoor <- dt[,.(timestamp,indoor_temperature)]
```

# Rolling Mean

```{r}
## Perform 14 Days Hourly Rolling Average

# Next, the hourly temperature is used to calculate the 14 days temperature backwards. 

# Invert table so slide window goes backwards
inverted_outdoor <- outdoor[order(-ymd_h)]

# Calculate rolling mean and reverts back the output from the rolling mean. 
rolling_mean_outdoor_temperature <- rev(rollapply(inverted_outdoor$outdoor_temperature,
                                            width=n_hours_in_every_step,
                                            FUN=mean))

outdoor <- outdoor[n_hours_in_every_step:.N]  
outdoor$rolling_mean_outdoor_temperature <- rolling_mean_outdoor_temperature
colnames(outdoor) <- c("timestamp","outdoor_temperature","rolling_mean_outdoor_temperature")
```

# Merge Outdoor and Indoor

```{r}
dt <- merge(outdoor,indoor,by="timestamp")

# Add Rolling Mean Vector to Simulation Data Column-Wise. Notice this is NOT an inner join.
#min_rows <- min(length(rolling_temperature_f_mean),nrow(simulation_data))

#dt <- simulation_data[1:min_rows]
#dt$rolling_temperature_f_mean <- rolling_temperature_f_mean[1:min_rows]
```


## Calculate boundaries

The comfort level boundaries are calculated using the outdoor rolling average adjusted by the indoor airspeed. 

```{r}
calculate_air_speed_adjustment <- function(in_air_speed){
  return(0.0153*in_air_speed+0.4333)
}

dt$upper_bound_0_fpm <- 0.31*dt$rolling_mean_outdoor_temperature + 60.5
dt$upper_bound_120_fpm <- dt$upper_bound_0_fpm + calculate_air_speed_adjustment(120)
dt$upper_bound_200_fpm <- dt$upper_bound_0_fpm + calculate_air_speed_adjustment(200)


dt$lower_bound <- 0.31*dt$rolling_mean_outdoor_temperature + 47.9 
```

## Calculate Acceptance

With the comfort upper and lower bound, we then compare the indoor temperature against them.

```{r}
dt$acceptable_temperature_80 <- NA_character_
dt[indoor_temperature > upper_bound_200_fpm]$acceptable_temperature_80 <- "Unacceptable Hot"
dt[indoor_temperature > upper_bound_120_fpm & indoor_temperature <= upper_bound_200_fpm]$acceptable_temperature_80 <- "Acceptable at 200 fpm"
dt[indoor_temperature > upper_bound_0_fpm & indoor_temperature <= upper_bound_120_fpm]$acceptable_temperature_80 <- "Acceptable at 120 fpm"
dt[indoor_temperature > lower_bound & indoor_temperature < upper_bound_0_fpm]$acceptable_temperature_80 <- "Acceptable at 0 fpm"
dt[indoor_temperature < lower_bound]$acceptable_temperature_80 <- "Unacceptable Cold"

```

## Calculate Degrees Off

In addition, how far higher or lower than the upper and lower bound respectively can also be calculated.

```{r}
dt$degrees_off_0_fpm <- NA_real_

dt[indoor_temperature > upper_bound_0_fpm]$degrees_off_0_fpm <- dt[indoor_temperature > upper_bound_0_fpm]$indoor_temperature - dt[indoor_temperature > upper_bound_0_fpm]$upper_bound_0_fpm

dt[indoor_temperature < lower_bound]$degrees_off_0_fpm <- dt[indoor_temperature < lower_bound]$indoor_temperature - dt[indoor_temperature < lower_bound]$lower_bound

dt[indoor_temperature > lower_bound & indoor_temperature < upper_bound_0_fpm]$degrees_off_0_fpm <- 0
```

# Output Table And Optional Time Slice

Finally, we the output table containing all the calculated information.

```{r Time Window Filter for Simulated Data}
kable(head(dt))


#dt <- dt[timestamp > ymd("1800-10-06") | timestamp < ymd("1800-06-22")] # New time window
p_start_time <- "1800-10-06"
p_end_time <- "1800-06-22"
dt <- dt[timestamp > ymd("1800-10-06") | timestamp < ymd("1800-02-09")] #Old Time Window
#p_end_time <- "1800-02-09"
```

# Plot

```{r fig.width= 9, fig.height=4}
n_acceptable <- nrow(dt[acceptable_temperature_80 == "Acceptable at 0 fpm"])
n_unacceptable <- nrow(dt[acceptable_temperature_80 != "Acceptable at 0 fpm"])

percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}

n_acceptable_percent <- percent(n_acceptable/(n_acceptable+n_unacceptable))
n_unacceptable_percent <- percent(n_unacceptable/(n_acceptable+n_unacceptable))

st <- min(dt$timestamp)
et <- max(dt$timestamp)
#p_start_time <- paste0(year(st),"-",month(st),"-",day(st))
#p_end_time <- paste0(year(et),"-",month(et),"-",day(et))


status <- paste0(p_start_time," to ",p_end_time," | Acceptable 0 fpm: ",n_acceptable," (",n_acceptable_percent,") ; Unacceptable 0 fpm: ",n_unacceptable," (",n_unacceptable_percent,")")



p <- ggplot(data = dt) + 
  geom_point(aes(rolling_mean_outdoor_temperature,indoor_temperature,color=acceptable_temperature_80)) + 
  geom_line(aes(rolling_mean_outdoor_temperature,upper_bound_200_fpm),color="black") + 
  geom_line(aes(rolling_mean_outdoor_temperature,upper_bound_120_fpm),color="black") + 
  geom_line(aes(rolling_mean_outdoor_temperature,upper_bound_0_fpm),color="black") + 
  geom_line(aes(rolling_mean_outdoor_temperature,lower_bound),color="black") + 
  
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$upper_bound_200_fpm), label = "fpm 200", vjust = -1), size = 3) + 
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$upper_bound_120_fpm), label = "fpm 120", vjust = -1), size = 3) + 
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$upper_bound_0_fpm)-0.6, label = "fpm 0", vjust = -1), size = 3) + 
  geom_text(aes(max(dt$rolling_mean_outdoor_temperature)+0.2,max(dt$lower_bound), label = "fpm 0", vjust = -1), size = 3) + 

  
  theme_minimal() + 
  scale_colour_manual(values=c("Unacceptable Cold"="#0072B2","Acceptable at 0 fpm"="#2e6d1e","Acceptable at 120 fpm"="#55bc3c","Acceptable at 200 fpm"="#3bfc0c","Unacceptable Hot"="#D55E00")) +
  ggtitle(paste0("Adaptive Method - ",data_type), subtitle = status) + 
  ylab("Indoor Temperature (F)") + 
  xlab("Rolling Mean Outdoor Temperature (F)") + 
  labs(color="Acceptable Temperature 80")

p
```


```{r}
counts <- rbind(
  data.table(Type="Unacceptable Hot",Count=nrow(dt[acceptable_temperature_80 == "Unacceptable Hot"]),Percent = nrow(dt[acceptable_temperature_80 == "Unacceptable Hot"])/nrow(dt)),
  data.table(Type="Acceptable at 200 fpm",Count=nrow(dt[acceptable_temperature_80 == "Acceptable at 200 fpm"]),Percent = nrow(dt[acceptable_temperature_80 == "Acceptable at 200 fpm"])/nrow(dt)),
  data.table(Type="Acceptable at 120 fpm",Count=nrow(dt[acceptable_temperature_80 == "Acceptable at 120 fpm"]),Percent = nrow(dt[acceptable_temperature_80 == "Acceptable at 120 fpm"])/nrow(dt)),
  data.table(Type="Acceptable at 0 fpm",Count=nrow(dt[acceptable_temperature_80 == "Acceptable at 0 fpm"]),Percent = nrow(dt[acceptable_temperature_80 == "Acceptable at 0 fpm"])/nrow(dt)),
  data.table(Type="Unacceptable Cold",Count=nrow(dt[acceptable_temperature_80 == "Unacceptable Cold"]),Percent = nrow(dt[acceptable_temperature_80 == "Unacceptable Cold"])/nrow(dt))
)
counts$Label <- "-"

counts <- counts[,.(Label,Type,Count,Percent)]

counts$cumsum <- counts$Count
counts$cumsum[3] <- counts$cumsum[4] + counts$cumsum[3] 
counts$cumsum[2] <- counts$cumsum[2] + counts$cumsum[3] 


counts$cumpercent <- counts$Percent
counts$cumpercent[3] <- counts$cumpercent[4] + counts$cumpercent[3] 
counts$cumpercent[2] <- counts$cumpercent[2] + counts$cumpercent[3] 

counts$Percent <- percent(counts$Percent)
counts$cumpercent <- percent(counts$cumpercent)

counts$Type2 <- c("Unacceptable Hot","Additional acceptable at 200 fpm","Additional acceptable at 120 fpm","Additional acceptable at 0 fpm","Unacceptable Cold")

counts <- counts[,.(Label,Type,cumsum,cumpercent,Type2,Count,Percent)]

formattable(counts,list(
            area(row=1,col = Label) ~ color_tile("#D55E00", "#D55E00"),
            area(row=2,col = Label) ~ color_tile("#a5ff00", "#a5ff00"),
            area(row=3,col = Label) ~ color_tile("#62d147", "#62d147"),
            area(row=4,col = Label) ~ color_tile("#2e6d1e", "#2e6d1e"),
            area(row=5,col = Label) ~ color_tile("#0072B2", "#0072B2")
            ))
```


```{r}
# 28.34 is pixel to cm, see: http://www.answers.com/Q/How_many_pixels_to_one_centimeter
ggsave(filename=save_path_and_name,plot=p,
     width=1340/(28.34*3),
     height=585/(28.34*3))
```


